---
title: "Plots for high dimensional data"
author: "Cory Whitney"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyverse) #tidyverse includes a number of useful packages
```

A number of useful plotting functions created using functions from `ggplot2`, `plyr`, `dplyr` and `tidyverse`.

##Radial bar plots

Create a data set for radial plots with a factor variable
```{r }
DF <- data.frame(variable = as.factor(1:10),
                 value = sample(10, replace = TRUE))
head(DF)
```

Create a function to compute standard error of mean to reprsent some of the uncertainty in the data
```{r }
se <- function(x) sqrt(var(x)/length(x)) 
set.seed(9876) 
```

Radial plot with a factor variable
```{r }
ggplot(DF, aes(variable, value, fill = variable)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_errorbar(aes(ymin = value - se(DF$value), 
                    ymax = value + se(DF$value), 
                    color = variable), 
                width = .2) + 
  scale_y_continuous(breaks = 0:nlevels(DF$variable)) +
  theme_minimal() +
  coord_polar()
```

###Radial bar plots with a number of factor variables

Create a data set for radial plots with with three factor variables
```{r }
DF2 <- data.frame(name = rep(letters[1:3], length.out = 30),
                 variable = as.factor(1:5),
                 variable2 = rep(letters[4:7], length.out = 30),
                 value = sample(10, replace = TRUE))
head(DF2)
```

Radial plots with three factor variables
```{r }
multi_plot <- ggplot(DF2, aes(variable, value, fill = variable)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_errorbar(aes(ymin = value - se(DF2$value), 
                    ymax = value + se(DF2$value), 
                    color = variable), 
                width = .2) + 
  scale_y_continuous(breaks = 0:nlevels(DF2$variable)) +
  theme_minimal() +
  coord_polar()
```

Rows as names and columns as variables `variable2`
```{r }
# Rows are name and columns are variable2
multi_plot + facet_grid(name ~ variable2)
```

Create a data for grid function with four factor variables and one continuous.
```{r }
DF3 <- data.frame(name = rep(letters[1:3], length.out = 600),
                  variable = as.factor(sample(5, replace = TRUE)),
                  variable2 = rep(letters[4:7], length.out = 600),
                  variable3 = rep(letters[8:16], length.out = 600),
                  value = sample(50, replace = TRUE))
head(DF3)
```

##Radial box plots

Radial box plot with a data set and grid with four factor variables and one continuous.
```{r }
multi_plot <- ggplot(data = DF3, aes(x=variable, y=value, fill=variable)) +
  geom_boxplot() +
  scale_y_continuous(breaks = 0:nlevels(DF3$variable)) +
  theme_minimal() +
  coord_polar()
multi_plot
```

Radial box plot with rows as names and columns as variables `variable2`
```{r }
multi_plot + facet_grid(name ~ variable2)
```

###Radial box plots example using `ToothGrowth` data 

```{r }
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
DF4 <- ToothGrowth
head(DF4)
```

```{r }
box_plot <- ggplot(DF4, aes(x=dose, y=len, group=dose)) + 
  geom_boxplot(aes(fill=dose)) +
  theme_minimal()+
  coord_polar()
```

Split the radial boxplot in a vertically

```{r }
# 
box_plot + facet_grid(supp ~ .)
```

Split the radial boxplot horizontally 
```{r }
box_plot + facet_grid(. ~ supp)
```

###Sunburst plot

Create data set for a sunburst-style barplot confined to a circle
```{r }
# make some fake data
DF5 <- data.frame(
  'level1'=c('a', 'a', 'a', 'a', 'b', 'b', 'c', 'c', 'c'), 
  'level2'=c('a1', 'a2', 'a3', 'a4', 'b1', 'b2', 'c1', 'c2', 'c3'), 
  'value'=c(.025, .05, .027, .005, .012, .014, .1, .03, .18))
head(DF5)
```

Create a sunburst-style barplot confined to a circle

```{r }
ggplot(DF5, aes(y=value)) +
  geom_bar(aes(fill=level1, x=0), width=.5, stat='identity') + 
  geom_bar(aes(fill=level2, x=.25), width=.25, stat='identity') + 
  coord_polar(theta='y') + 
  theme_minimal()
```

##Plot high dimensional data

Plot data which is all continuous

```{r }
DF6 <- data.frame(replicate(4,sample(1:200,1000,rep=TRUE)))
addme <- data.frame(replicate(1,sample(0:1,1000,rep=TRUE)))
DF6 <- cbind(DF6,addme)
colnames(DF6) <- c("var1","var2","var3","var4","var5")
```

###Bubble plot
```{r }
ggplot(DF6, aes(x=var1, y=var2, fill=var3, color=var4, size=var5)) +
  geom_point(shape=21) +
  #scale_color_gradient(low="grey", high="black") +
  #scale_color_gradientn(colours = rainbow(3))+
  scale_size_continuous(range=c(1,12))+ 
  scale_fill_gradient(low="grey", high="black")
```

Plot continuous and categorical data using `facet_grid`
Here we use `var3` and `var4` as quantiles to apply to plot area  

```{r }

DF6$var4.cat <- cut(DF6$var4, quantile(DF6$var4, (0:5)/5), include.lowest=T)
DF6$var3.cat <- cut(DF6$var3, quantile(DF6$var3, (0:5)/5), include.lowest=T)
```

```{r }
ggplot(DF6, aes(x=var1, y=var2, fill=var4, color=var3, size=var5^2)) +
  geom_point(shape=21) +
  scale_color_gradient(low="red", high="green") +
  scale_size_continuous(range=c(1,12)) +
  facet_grid(var3.cat ~ var4.cat)
```


