---
title: "Visualizing high dimensional data"
author: "Cory Whitney, Eduardo Fernandez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyverse) #tidyverse includes a number of useful packages
```

Dealing with large data sets 
Aims objectives and hypotheses should, of course, be determined before any data is collected 

Visualization is a powerful too to help get an overview of findings

It is, of course, also useful when reporting differences

Here wedemonstrate approaches to using visualization for several types of high dimensional data


A number of useful plotting functions created using functions from `ggplot2`, `plyr`, `dplyr` and `tidyverse`.

##Radial bar plots

Create a data set for radial plots with a factor variable
```{r }
DF <- data.frame(variable = as.factor(1:10),
                 value = sample(10, replace = TRUE))
head(DF)
```

Create a function to compute standard error of mean to reprsent some of the uncertainty in the data
```{r }
se <- function(x) sqrt(var(x)/length(x)) 
set.seed(9876) 
```

Radial plot with a factor variable
```{r }
ggplot(DF, aes(variable, value, fill = variable)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_errorbar(aes(ymin = value - se(DF$value), 
                    ymax = value + se(DF$value), 
                    color = variable), 
                width = .2) + 
  scale_y_continuous(breaks = 0:nlevels(DF$variable)) +
  theme_minimal() +
  coord_polar()
```

###Radial bar plots with a number of factor variables

Create a data set for radial plots with with three factor variables
```{r }
DF2 <- data.frame(name = rep(letters[1:3], length.out = 30),
                 variable = as.factor(1:5),
                 variable2 = rep(letters[4:7], length.out = 30),
                 value = sample(10, replace = TRUE))
head(DF2)
```

Radial plots with three factor variables
```{r }
multi_plot <- ggplot(DF2, aes(variable, value, fill = variable)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_errorbar(aes(ymin = value - se(DF2$value), 
                    ymax = value + se(DF2$value), 
                    color = variable), 
                width = .2) + 
  scale_y_continuous(breaks = 0:nlevels(DF2$variable)) +
  theme_minimal() +
  coord_polar()
```

Rows as names and columns as variables `variable2`
```{r }
# Rows are name and columns are variable2
multi_plot + facet_grid(name ~ variable2)
```

Create a data for grid function with four factor variables and one continuous.
```{r }
DF3 <- data.frame(name = rep(letters[1:3], length.out = 600),
                  variable = as.factor(sample(5, replace = TRUE)),
                  variable2 = rep(letters[4:7], length.out = 600),
                  variable3 = rep(letters[8:16], length.out = 600),
                  value = sample(50, replace = TRUE))
head(DF3)
```

##Radial box plots

Radial box plot with a data set and grid with four factor variables and one continuous.
```{r }
multi_plot <- ggplot(data = DF3, aes(x=variable, y=value, fill=variable)) +
  geom_boxplot() +
  scale_y_continuous(breaks = 0:nlevels(DF3$variable)) +
  theme_minimal() +
  coord_polar()
multi_plot
```

Radial box plot with rows as names and columns as variables `variable2`
```{r }
multi_plot + facet_grid(name ~ variable2)
```

###Radial box plots example using `ToothGrowth` data 

```{r }
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
DF4 <- ToothGrowth
head(DF4)
```

```{r }
box_plot <- ggplot(DF4, aes(x=dose, y=len, group=dose)) + 
  geom_boxplot(aes(fill=dose)) +
  theme_minimal()+
  coord_polar()
```

Split the radial boxplot in a vertically

```{r }
# 
box_plot + facet_grid(supp ~ .)
```

Split the radial boxplot horizontally 
```{r }
box_plot + facet_grid(. ~ supp)
```

###Sunburst plot

Create data set for a sunburst-style barplot confined to a circle
```{r }
# make some fake data
DF5 <- data.frame(
  'level1'=c('a', 'a', 'a', 'a', 'b', 'b', 'c', 'c', 'c'), 
  'level2'=c('a1', 'a2', 'a3', 'a4', 'b1', 'b2', 'c1', 'c2', 'c3'), 
  'value'=c(.025, .05, .027, .005, .012, .014, .1, .03, .18))
head(DF5)
```

Create a sunburst-style barplot confined to a circle

```{r }
ggplot(DF5, aes(y=value)) +
  geom_bar(aes(fill=level1, x=0), width=.5, stat='identity') + 
  geom_bar(aes(fill=level2, x=.25), width=.25, stat='identity') + 
  coord_polar(theta='y') + 
  theme_minimal()
```

###Spider plot

Create a function to obtain straight lines

```{r}
coord_radar <- 
  function(theta='x', start=0, direction=1){
    # input parameter sanity check
    match.arg(theta, c('x','y'))
    
    ggproto(
      NULL, CoordPolar, 
      theta=theta, r=ifelse(theta=='x','y','x'),
      start=start, direction=sign(direction),
      is_linear=function() TRUE)
  }
```

Create a new dataset and a spider plot

```{r}
factor <- c(rep("A", 16), rep("B", 16))
variable <- as.factor(c(1:16))
value <- sample(c(1:10),32,replace = T)
```

One way to close the plot is to add an "extra" level of variable which needs the same value as level 1. Both factors ("A" and "B") need this correction.

```{r}
value[16] <- value[1]
value[32] <- value[17]  
```

```{r}
data <- data.frame(factor = factor, variable = variable, value = value)
```

With `ggplot`

```{r}
ggplot(data, aes(as.numeric(data$variable),value)) + coord_radar() +
  geom_path() + scale_x_continuous(breaks = c(1:15)) +
  labs(x = "variable") + facet_wrap(.~ factor) + theme_minimal()
```

## Plot high dimensional data

Create data which is all continuous

```{r }
DF6 <- data.frame(replicate(4,sample(1:200,1000,rep=TRUE)))
addme <- data.frame(replicate(1,sample(0:1,1000,rep=TRUE)))
DF6 <- cbind(DF6,addme)
```

Add new column names

```{r }
colnames(DF6) <- c("var1","var2","var3","var4","var5")
```

### Bubble plot

```{r }
ggplot(DF6, aes(x=var1, y=var2, fill=var3, color=var4, size=var5)) +
  geom_point(shape=21) +
  #scale_color_gradient(low="grey", high="black") +
  #scale_color_gradientn(colours = rainbow(3))+
  scale_size_continuous(range=c(1,12))+ 
  scale_fill_gradient(low="grey", high="black")
```

## Plot continuous and categorical data


Here we use `var3` and `var4` as quantiles to apply to plot area  

```{r }

DF6$var4.cat <- cut(DF6$var4, quantile(DF6$var4, (0:5)/5), include.lowest=T)
DF6$var3.cat <- cut(DF6$var3, quantile(DF6$var3, (0:5)/5), include.lowest=T)
```

Plot continuous and categorical data using `facet_grid`

```{r }
ggplot(DF6, aes(x=var1, y=var2, fill=var4, color=var3, size=var5^2)) +
  geom_point(shape=21) +
  scale_color_gradient(low="red", high="green") +
  scale_size_continuous(range=c(1,12)) +
  facet_grid(var3.cat ~ var4.cat)
```
